---
title: "Practicum1"
output: html_notebook
---

```{r echo=FALSE}
# 1. Library
require(RMySQL)

# 2. Settings
db_user <- ''
db_password <- ''
db_name <- ''
db_host <- ''
db_port <- 3306
fpath = ""

# 3. Read data from db
mydb <-  dbConnect(MySQL(), user = db_user, password = db_password,
                 dbname = db_name, host = db_host, port = db_port)
```

#Create a set of SQL DDL statements for the above logical model and realize that schema in a new database.
```{sql connection=mydb}
DROP TABLE IF EXISTS Incident_Precipitation_Cond;
```
```{sql connection=mydb}
DROP TABLE IF EXISTS Precipitation_Condition;
```
```{sql connection=mydb}
DROP TABLE IF EXISTS Incident;
```
```{sql connection=mydb}
DROP TABLE IF EXISTS Airport;
```
```{sql connection=mydb}
DROP TABLE IF EXISTS Aircraft;
```
```{sql connection=mydb}
DROP TABLE IF EXISTS Wildlife;
```

```{sql connection=mydb}
CREATE TABLE Airport(
  Airport_ID INT NOT NULL,
  Airport_Name VARCHAR(255) NOT NULL,
  Origin_State VARCHAR(255) NOT NULL,
  PRIMARY KEY (Airport_ID)
);
```

```{sql connection=mydb}
CREATE TABLE Aircraft(
  Aircraft_ID INT NOT NULL,
  Make_model TEXT NOT NULL,
  Aircraft_Type VARCHAR(255) NOT NULL,
  Is_Large BOOLEAN NOT NULL,
  Number_Of_Engines INT NOT NULL,
  PRIMARY KEY (Aircraft_ID)
);
```

```{sql connection=mydb}
CREATE table Wildlife(
  Wildlife_ID INT NOT NULL,
  Size VARCHAR(255) NOT NULL,
  Species VARCHAR(255) NOT NULL,
  PRIMARY KEY (Wildlife_ID)
);
```

```{sql connection=mydb}
CREATE TABLE Incident (
  Record_ID INT NOT NULL,
  Impact_to_flight VARCHAR(255) NOT NULL,
  Flight_Date DATE NOT NULL,
  Is_damage_indicated BOOLEAN NOT NULL,
  Aircraft_ID INT NOT NULL,
  Wildlife_ID INT NOT NULL,
  Airport_ID INT NOT NULL,
  Actual_Number_Struck INT NOT NULL,
  Altitude BINARY NOT NULL,
  Number_of_injured INT NOT NULL,
  Remarks TEXT,
  Pilot_Warned BOOLEAN NOT NULL,
  Feet_above_ground INT NOT NULL,
  Cost INT NOT NULL,
  Phase VARCHAR(255) NOT NULL,
  Airline_Operator VARCHAR(255) NOT NULL,
  Remains_of_wildlife_collected BOOLEAN,
  Remains_of_wildlife_sent_to_Smithsonian BOOLEAN,
  PRIMARY KEY (Record_ID),
  FOREIGN KEY (Aircraft_ID) REFERENCES Aircraft(Aircraft_ID),
  FOREIGN KEY (Wildlife_ID) REFERENCES Wildlife(Wildlife_ID),
  FOREIGN KEY (Airport_ID) REFERENCES Airport(Airport_ID)
);
```

```{sql connection=mydb}
CREATE TABLE Precipitation_Condition(
  PC_ID INT NOT NULL,
  Conditions VARCHAR(255) NOT NULL,
  PRIMARY KEY (PC_ID)
);
```

```{sql connection=mydb}
CREATE TABLE Incident_Precipitation_Cond(
  Record_ID INT NOT NULL,
  PC_ID INT NOT NULL,
  PRIMARY KEY (Record_ID, PC_ID),
  FOREIGN KEY (Record_ID) REFERENCES Incident(Record_ID),
  FOREIGN KEY (PC_ID) REFERENCES Precipitation_Condition(PC_ID)
);
```




```{r}
df = read.csv('BirdStrikesData.csv')
```

#Omit the empty aircraft information
```{r}
df = df[which(df$Aircraft..Type!=""),]
```

#Rename each column name the same as the table attributes name
```{r}
colnames(df) = c("Record_ID","Aircraft_Type","Airport_Name","Altitude", "Make_model","Number_Struck","Actual_Number_Struck","Impact_To_Flight","Flight_Date",
                 "Is_Damage_Indicated",
 "Number_Of_Engines","Airline_Operator","Origin_State","Phase","Conditions",
 "Remains_of_wildlife_collected","Remains_of_wildlife_sent_to_Smithsonian","Remarks","Size",
 "Conditions_Sky","Species","Pilot_Warned","Cost","Feet_above_ground","Number_of_injured",
 "Is_Large")
```

#Create a column of airport id and extract columns to create a new airport dataframe.
#Delete duplicate rows in Airport dataframe
```{r}
#df$Airport_ID = cumsum(!duplicated(df$Airport_Name))
library(dplyr)
df$Airport_ID = df%>%group_indices(Airport_Name)
df_Airport = df[,c(27,3,13),drop=F]
df_Airport = df_Airport[!duplicated(df_Airport$Airport_ID),]
```

#Load the airport dataframe to airport table
```{r}
dbSendQuery(mydb, "SET GLOBAL local_infile = true;")
dbWriteTable(mydb, "Airport", df_Airport,overwite=TRUE,append=TRUE,row.names=F)
```

```{sql connection=mydb}
select * from Airport LIMIT 50;
```

#Create a column of aircraft id 
#Convert Is_Large attribute to boolean type
#Extract columns to create a new Aircraft dataframe
#Delete the duplicate rows in Aircraft dataframe
```{r}
df$Aircraft_ID = df %>% group_indices(Make_model)
df$Is_Large[df$Is_Large=="Yes"] <- 1
df$Is_Large[df$Is_Large=="No"] <- 0
df_Aircraft = df[,c(28,5,2,26,11),drop=F]
df_Aircraft = df_Aircraft[!duplicated(df_Aircraft$Aircraft_ID),]
```

#Load the Aircraft dataframe to Aircraft table
```{r}
dbSendQuery(mydb, "SET GLOBAL local_infile = true;")
dbWriteTable(mydb, "Aircraft", df_Aircraft,append=T,row.names = F)
```
```{sql connection=mydb}
select* from Aircraft LIMIT 50;
```



#Create a new column Wildlife_ID and extract columns to create a new Wildlife dataframe
Delete duplicate rows in Wildlife dataframe
```{r}
df$Wildlife_ID = df %>% group_indices(Species)
df_Wildlife = df[,c(29,19,21),drop=F]
df_Wildlife= df_Wildlife[!duplicated(df_Wildlife$Wildlife_ID),]
```

#Load Wildlife dataframe to Wildlife table
```{r}
dbSendQuery(mydb, "SET GLOBAL local_infile = true;")
dbWriteTable(mydb, "Wildlife", df_Wildlife,append=T,row.names = F)
```

```{sql connection=mydb}
select * from Wildlife LIMIT 50;
```

#Create a new column PC_ID and extract two columns to form a new dataframe
```{r}
df$PC_ID = df %>% group_indices(Conditions)
df_Incident_Precipitation_Cond = df[,c(1,30),drop=F]
```

#Load this dataframe Incident_Precipitation_Cond table
```{r}

dbWriteTable(mydb, "Incident_Precipitation_Cond",   
             df_Incident_Precipitation_Cond,overwrite = T,row.names = F)
```

```{sql connection=mydb}
select * from Incident_Precipitation_Cond LIMIT 50;
```
#Create a new dataframe called Precipitation_Condition based on Conditions and PC_ID
#Delete duplicate rows
```{r}
df_Precipitation_Condition = df[,c(30,15),drop=F]
df_Precipitation_Condition=  
  df_Precipitation_Condition[!duplicated(df_Precipitation_Condition$PC_ID),]
```

#Load Precipitation_Condition dataframe to Precipitation_Condition table
```{r}
dbWriteTable(mydb, "Precipitation_Condition",   
             df_Precipitation_Condition,append=T,row.names = F)
```
```{sql connection=mydb}
select * from Precipitation_Condition
```

#Convert Is_Damage_Indicated to boolean
```{r}
df$Is_Damage_Indicated[df$Is_Damage_Indicated=="Caused damage"] <- 1
df$Is_Damage_Indicated[df$Is_Damage_Indicated=="No damage"] <- 0
```

#Convert Pilot_Warned to boolean
```{r}
df$Pilot_Warned[df$Pilot_Warned=="Y"] <- 1
df$Pilot_Warned[df$Pilot_Warned=="N"] <- 0
```

#Convert Remains_of_wildlife_collected and Remains_of_wildlife_sent_to_Smithsonian to boolean
```{r}
df$Remains_of_wildlife_collected[df$Remains_of_wildlife_collected=="TRUE"] <- 1
df$Remains_of_wildlife_sent_to_Smithsonian[df$Remains_of_wildlife_sent_to_Smithsonian
                                           =="True"] <-1
```

#Change the Flight_Date format to %Y-%m-%d
```{r}
df$Flight_Date = gsub(" 0:00","",df$Flight_Date)
df$Flight_Date = strptime(as.character(df$Flight_Date),"%m/%d/%Y")
df$Flight_Date = format(as.Date(df$Flight_Date), "%Y-%m-%d")
```

#Extract columns to create a Incident dataframe
```{r}
df_Incident= df[,c(1,7,8,9,10,28,29,27,4,25,18,22,24,23,14,12,16,17),drop=F]
```

#Load the Incident dataframe to Incident table
```{r}
dbWriteTable(mydb, "Incident",   
             df_Incident,append=T,row.names = F)
```

```{sql connection=mydb}
select * from Incident LIMIT 50;
```
#Create a SQL query against your database to find the number of bird strike incidents for each airline upon take-off or climb. Include all airlines.
```{sql connection=mydb}
SELECT Airline_Operator, COUNT(Record_ID) AS number_of_incidents_takeoff_or_climb
FROM Incident
WHERE Phase = 'Climb' OR Phase = 'Take-off run'
GROUP BY Airline_Operator
```

#Create a SQL query against your database to find the airports that had the most bird strike incidents (during any flight phase). Include all airlines.
```{sql connection=mydb}
SELECT t2.Airport_Name
FROM Incident AS t1
LEFT JOIN Airport AS t2 ON t1.Airport_ID = t2.Airport_ID
GROUP BY t1.Airport_ID
HAVING COUNT(t1.Airport_ID) = (
  SELECT COUNT(Airport_ID)
  FROM Incident
  GROUP BY Airport_ID
  ORDER BY COUNT(Airport_ID) DESC
  LIMIT 1
)
```

#Create a SQL query against your database to find the number of bird strike incidents by year. Include all airlines
```{sql connection=mydb}
SELECT EXTRACT(year FROM Flight_Date) AS year, COUNT(Record_ID) AS number_of_incidents
FROM Incident
GROUP BY year
ORDER BY year
```

#Visualizes the number of bird strikes incidents per year from 2008 to 2011 during take-off/climbing
```{r}
library(ggplot2)

sql <- "SELECT year, number_of_incidents FROM (SELECT EXTRACT(year FROM Flight_Date) AS year, COUNT(Record_ID) AS    number_of_incidents
      FROM Incident
      WHERE Phase = 'Climb' OR Phase = 'Take-off run'
      GROUP BY year
      ORDER BY year) year_incident
      WHERE year >= 2008"
df <-dbGetQuery(mydb,sql)

ggplot(df, aes(x = year, y = number_of_incidents)) +
  geom_bar(stat='identity',fill = "#97B3C6") +
  theme_bw() +
  labs(x = "Year", y = "Number of Incidents", title = "strikes incidents per year during take-off/climbing")

```
#Visualizes the number of bird strikes incidents per year from 2008 to 2011 during descent/approach/landing
```{r}
sql <- "SELECT year, number_of_incidents FROM (SELECT EXTRACT(year FROM Flight_Date) AS year, COUNT(Record_ID) AS    number_of_incidents
      FROM Incident
      WHERE Phase = 'Approach' OR Phase = 'Landing Roll' OR Phase = 'Descent'
      GROUP BY year
      ORDER BY year) year_incident
      WHERE year >= 2008"
df <-dbGetQuery(mydb,sql)

ggplot(df, aes(x = year, y = number_of_incidents)) +
  geom_bar(stat='identity',fill = "#97B3C6") +
  theme_bw() +
  labs(x = "Year", y = "Number of Incidents", title = "strikes incidents per year during descent/approach/landing")
```

#Create a stored procedure that removes a bird strike incident from the database
```{sql connection=mydb}
DROP PROCEDURE IF EXISTS remove_bird_strike;
```
```{sql connection=mydb}
CREATE PROCEDURE remove_bird_strike (IN p_record_id INT)
BEGIN
  DELETE FROM Incident
  WHERE Record_ID = p_record_id;
END;
```

Before procedure call:
```{sql connection=mydb}
SELECT * FROM Incident where Record_ID = 1195;
```

Call procedure:
```{sql connection=mydb}
CALL remove_bird_strike(1195);
```

After procedure call:
```{sql connection=mydb}
SELECT * FROM Incident where Record_ID = 1195;
```


```{r}
dbDisconnect(mydb)
```